---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

devtools::load_all(export_all = FALSE)
```

# box.shiny

<!-- badges: start -->
<!-- badges: end -->

Create modularized `{shiny}` apps using the [`{box}`](https://klmr.me/box/index.html) package.
Contains further functions to facilitate the development of your shiny apps.


## Installation

So far, I have developed this package for my personal usage mainly.
However, I am very happy if you find the package interesting and are eager to use it.
In that case you can download it from Github:

``` r
# install.packages("remotes")
remotes::install_github("DavidJesse21/box.shiny")
```

Since the functions of this package do not import any functions from the `{box}` package 
you need to download that one as well.
You can install it from CRAN. However, for using it with this package I'd currently advise you 
to download this specific version of the package:

```r
install.packages('https://github.com/klmr/box/files/7204838/box_1.1.9000.tar.gz', type = 'source', repos = NULL)
```

This version contains the function `box::enable_autoreload()` which you will probably want to use such that 
modifications to your module files will immediately be adopted.


## Motivation

When developing a shiny app it is good practice to make use of [shiny modules](https://shiny.rstudio.com/articles/modules.html).
The linked article also gives you some advice on how to organize your modules in your app project, such as creating an `/R` directory where you can place your modules.
This is also similar to the way modules are organized when using the [`{golem}`](https://github.com/ThinkR-open/golem) framework for creating [production grade shiny apps](https://engineering-shiny.org/) where your app is organized as an R package.

Creating your app as an R package is a handy solution on the one hand as you can make use of the many features of R packages 
as well as of other packages facilitating package development.<br>
On the other hand, it can also be a factor that I consider to be restricting as you need to put all your `.R` files in the 
same `/R` directory.
The only way to organize your R-files within this directory is to use proper naming conventions, but you can't 
structure them by using additional directories.
For apps that are not too large, this may work fine. However, when working on larger shiny projects it can become hard 
to keep a good overview, especially when collaborating with other colleagues.

As an example consider a shiny app. It could e.g. be a navbar page or a different layout in which you can 
include different tabs.
Inside of those tabs there might again be different components that can be separated from each other:

```{r echo=FALSE}
fs::dir_create("app")
fs::dir_create("app", "tab1")
fs::dir_create("app", "tab1", "plot")
fs::dir_create("app", "tab1", "table")
fs::dir_create("app", "tab1", "aaa")
fs::dir_create("app", "tab2")
fs::dir_create("app", "tab2", "input-panel")
fs::dir_create("app", "tab2", "bbb")
fs::dir_create("app", "tab3")

fs::dir_tree("./app")

fs::dir_delete("app")
```


You can directly see the nested structure of the app.
`{box.shiny}` aims to adapt to that by supporting you to create shiny modules while using the [`{box}`](https://github.com/klmr/box) package.





## Usage

As already mentioned, before using this package you first need to install the `{box}` package and you should also get familiar 
with its syntax and functionality by browsing through its documentation.

You can then start your project using `{box.shiny}` with:

```{r}
# box.shiny::setup_project("name-of-directory-for-shiny-modules")
box.shiny::setup_project("shinymodules")
```

In the output of this function call you can see, what this function has done:
It has created two files in your root directory: `app.R` is the file that needs to be run in order to launch your application.
`shinymodules_dir.txt` is a simple text file that contains the relative path to the directory in which your shiny modules 
will be placed. This file should (usually) not be edited or deleted.

Furthermore, you can see that an `/app` directory has been created with some further files and directories.
The `app_server.R`, `app_ui.R` and `run_app.R` files are quite self-explanatory.
The latter simply contains a wrapper function around `shiny::shinyApp()` and gets imported in `app.R`.

The most important directory is `app/shinymodules`.
By using the functions `box.shiny::add_module_dir(...)` and/or `box.shiny::add_module_file(...)` you can add 
new files and/or directories to it.
The syntax is just similar to `fs::path(...)`.

```{r}
suppressMessages({
  box.shiny::add_module_file("main", open = FALSE)
  box.shiny::add_module_dir("tabs")
  box.shiny::add_module_file("tabs", "first-tab", open = FALSE)
})

fs::dir_tree("./app/shinymodules")
```

The files created with `add_module_file()` come with a template for your shiny module.
On the one hand you have the module's function definitions for the UI and server part.
On the other hand it also comes with an import statement that you need to copy into the 
`box::use()` declaration as well as the code for the function calls for the module.


You can add both server and UI helper/utility functions to `/app/helpers/ui` and `/app/helpers/server` with the
`add_helper_` functions:

```{r eval=FALSE}
box.shiny::add_helper_server("server-utils")
box.shiny::add_helper_ui("buttons")
```


For shiny apps I find it very useful to use R6 class objects. That is why there is the `/app/R6` directory.
You can add a new file containing a minimal template to it using:

```{r eval=FALSE}
box.shiny::add_R6Class("my-class")
```



Finally, there is also an `app/www/styles` directory with a `main.scss` file and a `/partials` directory in it.
You can add sass partials to the latter directory with:

```{r eval=FALSE}
box.shiny::add_sass_partial("first-stylesheet")
```

By default, this command will also modify `main.scss` by adding an import statement for the newly added sass partial
to it.



For a further and more detailed documentation see the vignettes [TODO].



## Acknowledgment

First and foremost thank you to  [Konrald Rudolph](https://github.com/klmr) for creating the `{box}` package, 
on which the idea for this package is based.

Apart from that, I have been using the `{golem}` package by the [ThinkR](https://github.com/ThinkR-open) team before a lot which has
not only been a main inspiration for this package but also a source of how I can implement my ideas in code.
E.g. the `add_module_` functions from this package have been written very similarly to golem's 
[`add_module`](https://thinkr-open.github.io/golem/reference/add_module.html) function.



```{r include=FALSE}
fs::dir_delete("./app")
fs::file_delete("app.R")
fs::file_delete("shinymodules_dir.txt")
```

